---
title: "Hands-on Exercise 1"
author: "Cheryl Jeanne Chiew"
date: "`r Sys.Date()`"
---

# Getting Started

## Installing and launching R packages

The code chunk below uses p_load() of pacman package to check if tidyverse packages are installed in the computer. If they are, then they will be launched into R.

```{r}
pacman::p_load(tidyverse)
```

## Importing the data

```{r}
exam_data <- read_csv("data/Exam_data.csv")
```

# Plotting a simple bar chart

```{r}
ggplot(data=exam_data, 
       aes(x=RACE)) +
  geom_bar()
```

# Plotting a trellis boxplot (Slide 43)

```{r}
ggplot(data=exam_data, 
       aes(y = MATHS, x= CLASS)) +
  geom_boxplot() +
  facet_grid(~ GENDER)
```

# Plotting a trellis boxplot (Slide 44)

```{r}
ggplot(data=exam_data, 
       aes(y = MATHS, x= CLASS)) +
  geom_boxplot() +
  facet_grid(rows = vars(GENDER))
```

# Plotting a trellis boxplot (Slide 45)

```{r}
ggplot(data=exam_data, 
       aes(y = MATHS, x= CLASS)) +
  geom_boxplot() +
  facet_grid(vars(GENDER), vars(CLASS))
```

# Working with theme (Slide 53)

```{r}
ggplot(data=exam_data, 
       aes(x=RACE)) +
  geom_bar() +
  coord_flip() +
  theme_minimal() +
  theme(panel.background = element_rect(
    fill = "lightblue",
    colour = "lightblue",
    linewidth = 0.5,
    linetype = "solid"),
    panel.grid.major = element_line(
      linewidth = 0.5,
      linetype = 'solid',
      colour = "white"), 
    panel.grid.minor = element_line(
      linewidth = 0.25,
      linetype = 'solid',
      colour = "white"))
```

# Designing Data Driven Graphics for Analysis I (Slide 56)

```{r}
exam_data %>%
  mutate(RACE = fct_infreq(RACE)) %>%
  ggplot(aes(x = RACE)) + 
  geom_bar()+
  ylim(0,220) +
  geom_text(stat="count", 
      aes(label=paste0(..count.., ", ", 
      round(..count../sum(..count..)*100,
            1), "%")),
      vjust=-1) +
  xlab("Race") +
  ylab("No. of\nPupils") +
  theme(axis.title.y=element_text(angle = 0))
```

# Designing Data Driven Graphics for Analysis II (Slide 59)

```{r}
ggplot(data=exam_data, 
       aes(x= MATHS)) +
  geom_histogram(bins=20, 
                 fill = "lightblue", 
                 color = "black") +
  geom_vline(aes(xintercept=mean(MATHS)), 
             color="red", 
             linetype="dashed") + 
  geom_vline(aes(xintercept=median(MATHS)), 
             color="black", 
             linetype="dashed")
```

# Designing Data Driven Graphics for Analysis III (Slide 60)

```{r}
d <- exam_data   
d_bg <- d[, -3]  

ggplot(d, aes(x = ENGLISH, fill = GENDER)) +
  geom_histogram(data = d_bg, fill = "grey", alpha = .5) +
  geom_histogram(colour = "black") +
  facet_wrap(~ GENDER) +
  guides(fill = FALSE) +  
  theme_bw()
```

# Designing Data Driven Graphics for Analysis IV (Slide 62)

```{r}
ggplot(data=exam_data, 
       aes(x= MATHS, y=ENGLISH)) +
  geom_point() +
  coord_cartesian(xlim=c(0,100),
                  ylim=c(0,100)) +
  geom_vline(aes(xintercept=50), 
             color="grey", 
             linetype="dashed", 
             linewidth = 1) +
  geom_hline(aes(yintercept=50), 
             color="grey", 
             linetype="dashed", 
             linewidth = 1) 

```
