---
title: "Hands On Exercise 05"
author: "Cheryl Jeanne Chiew"
date: "12 February 2023"
date-modified: "`r Sys.Date()`"
execute:
  echo: true
  eval: true
  warning: false
format: html
editor: visual
---

## Building Ternary Plots

### Installing and Loading R Packages

```{r}
pacman::p_load(ggtern, plotly, tidyverse)
```

### Importing Data

```{r}
pop_data <- read_csv("data/respopagsex2000to2018_tidy.csv") 
```

### Data Preparation

```{r}
agpop_mutated <- pop_data %>%
  mutate(`Year` = as.character(Year))%>%          #<< change datatype of YEAR to character
  pivot_wider(names_from = AG, 
              values_from = Population) %>%       #<< transpose values from AG into columns
  mutate(YOUNG = rowSums(.[4:8]))%>%              #<< sum of values by rows
  mutate(ACTIVE = rowSums(.[9:16]))  %>%
  mutate(OLD = rowSums(.[17:21])) %>%
  mutate(TOTAL = rowSums(.[22:24])) %>%
  filter(Year == 2018)%>%                         #<< filter for Year 2018 only
  filter(TOTAL > 0)                               #<< filter where total count > 0
```

### Ternary Plot

```{r}
ggtern::ggtern(data=agpop_mutated, aes(x=YOUNG,y=ACTIVE, z=OLD)) +
  geom_point() +
  labs(title="Population structure, 2015") +
  theme_rgbw()
```

### Interactive Ternary Plot

```{r}
label <- function(txt) {
  list(
    text = txt, 
    x = 0.1, y = 1,
    ax = 0, ay = 0,
    xref = "paper", yref = "paper", 
    align = "center",
    font = list(family = "serif", size = 15, color = "white"),
    bgcolor = "#b3b3b3", bordercolor = "black", borderwidth = 2
  )
}

axis <- function(txt) {
  list(
    title = txt, tickformat = ".0%", tickfont = list(size = 10)
  )
}

ternaryAxes <- list(
  aaxis = axis("Young"), 
  baxis = axis("Active"), 
  caxis = axis("Old")
)

plot_ly(
  agpop_mutated, 
  a = ~YOUNG, 
  b = ~ACTIVE, 
  c = ~OLD, 
  color = I("black"), 
  type = "scatterternary"
) %>%
  layout(
    annotations = label("Ternary Markers"), 
    ternary = ternaryAxes
  )
```



## Correlation Matrices

### Installing and Loading R Packages

```{r}
pacman::p_load(corrplot, ggstatsplot, tidyverse)
```

### Importing Data

```{r}
wine <- read_csv("data/wine_quality.csv")
```

### Basic Correlation Matrix

```{r}
pairs(wine[,1:11])
```

### Including Plot with Correlation Coefficients

```{r}
panel.cor <- function(x, y, digits=2, prefix="", cex.cor, ...) {
usr <- par("usr")
on.exit(par(usr))
par(usr = c(0, 1, 0, 1))
r <- abs(cor(x, y, use="complete.obs"))
txt <- format(c(r, 0.123456789), digits=digits)[1]
txt <- paste(prefix, txt, sep="")
if(missing(cex.cor)) cex.cor <- 0.8/strwidth(txt)
text(0.5, 0.5, txt, cex = cex.cor * (1 + r) / 2)
}

pairs(wine[,2:12], 
      upper.panel = panel.cor)
```

### Plot using ggCorrmat 

```{r}
grouped_ggcorrmat(
  data = wine,
  cor.vars = 1:11,
  grouping.var = type,
  type = "robust",
  p.adjust.method = "holm",
  plotgrid.args = list(ncol = 2),
  ggcorrplot.args = list(outline.color = "black", 
                         hc.order = TRUE,
                         tl.cex = 10),
  annotation.args = list(
    tag_levels = "a",
    title = "Correlogram for wine dataset",
    subtitle = "The measures are: alcohol, sulphates, fixed acidity, citric acid, chlorides, residual sugar, density, free sulfur dioxide and volatile acidity",
    caption = "Dataset: UCI Machine Learning Repository"
  )
)
```


### Visualising using Corrplot Package

```{r}
wine.cor <- cor(wine[, 1:11])

corrplot(wine.cor, 
         method = "ellipse",     #<< appear as semicircle rather than full circle
         type="lower",           #<< only show the lower half
         diag = FALSE,           #<< turn of diagonal cells
         tl.col = "black")       #<< axis labels to black
```

### Including numerical data in matrix

```{r}
corrplot.mixed(wine.cor, 
               lower = "ellipse",     #<< appear as semicircle rather than full circle
               upper = "number",      #<< show numbers in upper side
               tl.pos = "lt",
               diag = "l",
               tl.col = "black")
```


### Including test of significance

```{r}
wine.sig = cor.mtest(wine.cor, conf.level= .95)      #<< compute the p-values and confidence interval

corrplot(wine.cor,
         method = "number",
         type = "lower",
         diag = FALSE,
         tl.col = "black",
         tl.srt = 45,
         p.mat = wine.sig$p,
         sig.level = .05)


```

### Reordering using hierarchical clustering

```{r}
corrplot(wine.cor, 
         method = "ellipse", 
         tl.pos = "lt",
         tl.col = "black",
         order="hclust",
         hclust.method = "ward.D",       #<< clustering method
         addrect = 3)                    #<< 3 clusters defined
```


## Heatmaps for Multivariate Data

### Installing and Loading R Packages

```{r}
pacman::p_load(seriation, dendextend, heatmaply, tidyverse)
```

### Importing Data

```{r}
wh <- read_csv("data/WHData-2018.csv")
```

### Data Preparation

```{r}
row.names(wh) <- wh$Country                   #<< change the rows by country name instead of row number

wh1 <- dplyr::select(wh, c(3, 7:12))          #<< select interested column values

wh_matrix <- data.matrix(wh)                  #<< transform into matrix
```


### Heatmap() of R Stats

```{r}
wh_heatmap <- heatmap(wh_matrix,
                      scale="column",          #<< Scaling of values
                      cexRow = 0.6,            #<< used to define the font size used for y-axis and x-axis labels 
                      cexCol = 0.8,
                      margins = c(10, 4))      #<< ensure that the entire x-axis labels are displayed completely
```


### Interactive Heatmap using HeatMaply

```{r}
heatmaply(wh_matrix[, -c(1, 2, 4, 5)],
          scale = "column")                    #<< Scaling of values - when assumed normal distribution
```

```{r}
heatmaply(normalize(wh_matrix[, -c(1, 2, 4, 5)]))
#<< normalization is used when distribution is non-normal
```


### Visualizing Hierarchical Clustering 

```{r}
heatmaply(normalize(wh_matrix[, -c(1, 2, 4, 5)]),
          dist_method = "euclidean",
          hclust_method = "ward.D")
```


### Statistical Approach to Hierarchical Clustering

Determining the best method

```{r}
wh_d <- dist(normalize(wh_matrix[, -c(1, 2, 4, 5)]), method = "euclidean")
dend_expend(wh_d)[[3]]               #<< used to determine the recommended clustering method to be used - USE THE METHOD WITH THE HIGHEST OPTIMUM VALUE
```

Determining the optimal number of cluster

```{r}
wh_clust <- hclust(wh_d, method = "average")     #<< input based on best method
num_k <- find_k(wh_clust)
plot(num_k)
```

Cluster based on Optimum Inputs

```{r}
heatmaply(normalize(wh_matrix[, -c(1, 2, 4, 5)]),
          dist_method = "euclidean",
          hclust_method = "average",
          k_row = 3)
```


### Optimal Leaf Ordering (OLO) - Alt Method for Ordering Constraints due to HClustering

```{r}
heatmaply(normalize(wh_matrix[, -c(1, 2, 4, 5)]),
          seriate = "GW")                     #<< seriate = OLO or GW
```


### Working with Colour Palettes

```{r}
heatmaply(normalize(wh_matrix[, -c(1, 2, 4, 5)]),
          Colv=NA,
          seriate = "none",
          colors = Blues,                         #<< heatmap colour themes
          k_row = 5,                              #<< 5 clusters
          margins = c(NA,200,60,NA),              #<< change the top margin to 60 and row margin to 200   
          fontsize_row = 4,                       #<< change the font size for row and column labels
          fontsize_col = 5,
          main="World Happiness Score and Variables by Country, 2018 \nDataTransformation using Normalise Method",
          xlab = "World Happiness Indicators",
          ylab = "World Countries"
          )
```


## Parallel Coordinates Plots

### Installing and Loading R Packages

```{r}
pacman::p_load(GGally, parallelPlot, tidyverse)
```

### Importing Data

```{r}
wh <- read_csv("data/WHData-2018.csv")
```

### Plotting Static Parallel Coordinates Plot

```{r}
ggparcoord(data = wh, 
           columns = c(7:12), 
           groupColumn = 2,
           scale = "uniminmax",
           alphaLines = 0.2,
           boxplot = TRUE, 
           title = "Parallel Coordinates Plot of World Happines Variables")
```