---
title: "Take Home Exercise 04"
author: "Cheryl Jeanne Chiew"
date: "17 February 2023"
date-modified: "`r Sys.Date()`"
execute:
  echo: true
  eval: true
  warning: false
format: html
editor: visual
---

# 1. Introduction

In this take-home exercise, you are required to uncover the impact of COVID-19 as well as the global economic and political dynamic in 2022 on Singapore bi-lateral trade (i.e. Import, Export and Trade Balance) by using appropriate analytical visualisation techniques learned in Lesson 6: It’s About Time. Students are encouraged to apply appropriate interactive techniques to enhance user and data discovery experiences.

The dataset selected for this exercise is sourced from [Singstat.gov.sg](https://www.singstat.gov.sg/find-data/search-by-theme/trade-and-investment/merchandise-trade/latest-data) and contains data on the *Merchandise Imports and Exports by Region / Market*.

For the purpose of this exercise, the study period will be conducted from January 2020 to December 2022.


# 2. Data Visualization

#### Installing and Loading R Packages

```{r}
pacman::p_load(readxl, dplyr, tidyverse) 
```

#### Importing Data

```{r}
#| warning: false
import_data <- read_xlsx("data/outputFile.xlsx",
                         sheet = "T1",                    #<< select sheet T2
                         range = "A10:UT129",             #<< only read specified cell range 
                         na = "na")                       #<< treat na as missing values

export_data <- read_xlsx("data/outputFile.xlsx",
                         sheet = "T2",                    #<< select sheet T2
                         range = "A10:UT101",             #<< only read specified cell range
                         na = "na")                       #<< treat na as missing values
```

#### Data Preparation

```{r}
#| warning: false
import_data

##remove columns to only retain January 2020 to December 2022
##remove total and region rows
##Pivot table to put dates as row names
##transform row names to MM-YYYY
##Retain all countries, pick out nuances

##consider using horizon graph? for countries
##cycle plot
##slope graphs

```
